---
title: "Study 1"
author: "Brendan Kennedy"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, 
                    encoding = encoding, 
                    output_file=file.path("output", "markdown", "study1.html"))})
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r init, echo=FALSE, message=FALSE, warning=FALSE}

library(here)
library(tidyverse)
library(reticulate)
library(lme4)
library(lmerTest)

LV_ROOT <- Sys.getenv("LV_ROOT")
VENV_PATH <- Sys.getenv("VENV_PATH")
ROUND <- "round1"

options(qwraps2_markup = "markdown")

use_virtualenv(VENV_PATH, required = TRUE)
use_python(file.path(VENV_PATH, "bin", "python3"), required = TRUE)

source(here("src", "R", "study1.R"))
source(here("src", "R", "plotting.R"))

```

```{r data_loading, echo=FALSE, message=FALSE}

data_list <- load_data(data_round = ROUND)
head(data_list$participants, n=5)

```

The following table displays aggregate demographic statistics for the sample ($N$ = 208) following the exclusion of participants ($n$=10) falling in small clusters (those containing less than 5 participants).


```{r summarize, echo=FALSE, results="asis"}
summarize_demographics(data_list$participants)

```

The same statistics on the 10 participants that were dropped:

```{r summarize_dropped, echo=FALSE, results="asis"}

summarize_demographics(data_list$dropped.participants)

```

Summary statistics for our dependent variables, grouped by subject pools (USC and Community):

```{r dv_summary, echo=FALSE, results="asis"}
summarize_spin_and_span(data_list$participants)
```

## Vizualize Cluster Demographics

```{r cluster_demo_viz, echo=F, warning=F, message=F}

plt_df <- preprocess_participants_for_plotting(data_list$participants)
bar_plots <- plot_demo_as_bar_charts(plt_df, output_dir=here("output", "figures", ROUND, "clusters"))

```

## Inter-cluster Similarities (ARI)

How similar are cluster method outputs to each other? If we treat a categorical demographic variable as a ground truth cluster, what are the observed similarities? We answer these questions using the adjusted Rand Index, which approximates the probability of a pair of observations within the first clustering also co-occurring in the second.

```{r cluster_similarity, echo=F}

source_python(here("src", "python", "clustering.py"))

ari_mat <- compute_ari_stats(participant_df = data_list$participants, eval_type = 'num1')  # one sided ARI
ari_plot <- plot_ari(ari_mat)
ggsave(here("output", "figures", ROUND, "ari_barchart.png"), width=4, height=2, units='in')
ari_plot

```

## Are Clusters Predictive of Performance? 

```{r cluster_predictiveness, echo=F, results='asis', warning=F, message=F}

cluster_performance_results <- run_spin_models(data_list$participants)

knitr::kable(cluster_performance_results$spin_high_table)
#print(xtable::xtable(cluster_performance_results$spin_high_table, type='latex'), 
      #sanitize.text.function = function(x) {x},
      #include.rownames=F)
#print(xtable::xtable(cluster_performance_results$spin_low_table, type='html'), 
      #sanitize.text.function = function(x) {x},
      #include.rownames=F)
#print(xtable::xtable(cluster_performance_results$span_table, type='html'), 
      #sanitize.text.function = function(x) {x},
      #include.rownames=F)

res <- cluster_performance_results
  emm.spin.h <- emmeans::emmeans(res$spin.h.cluster, "cluster", type='response') %>%
    as.data.frame() %>%
    rename(response=rate) %>%
    add_column(task="SPiN-High")
  emm.spin.l <- emmeans::emmeans(res$spin.l.cluster, "cluster", type='response') %>%
    as.data.frame() %>%
    rename(response=rate) %>%
    add_column(task="SPiN-Low")
  emm.span <- emmeans::emmeans(res$span.cluster, "cluster", type='response') %>%
    as.data.frame() %>%
    add_column(task="SPAN")
  
spin_lims <- c(min(min(emm.spin.h$asymp.LCL), min(emm.spin.l$asymp.LCL)),
               max(max(emm.spin.h$asymp.UCL), max(emm.spin.l$asymp.UCL)))

p_1 <- ggplot(emm.spin.h, aes(x=reorder(cluster, response), y=response)) + 
    geom_point(shape=7) +
    geom_errorbar(aes(ymin=asymp.LCL,
                      ymax=asymp.UCL),
                  width=.4) +
    jtools::theme_apa(remove.y.gridlines = F, remove.x.gridlines = F) +
    xlab("Cluster") + ylab("SPiN-High EMM") +
    ylim(spin_lims)
p_2 <- ggplot(emm.spin.l, aes(x=reorder(cluster, response), y=response)) + 
    geom_point(shape=7) +
    geom_errorbar(aes(ymin=asymp.LCL,
                      ymax=asymp.UCL),
                  width=.4) +
    jtools::theme_apa(remove.y.gridlines = F, remove.x.gridlines = F) +
    xlab("Cluster") + ylab("SPiN-Low EMM") + 
    ylim(spin_lims)
p_3 <- ggplot(emm.span, aes(x=reorder(cluster, response), y=response)) + 
    geom_point(shape=7) +
    geom_errorbar(aes(ymin=asymp.LCL,
                      ymax=asymp.UCL),
                  width=.4) +
    jtools::theme_apa(remove.y.gridlines = F, remove.x.gridlines = F) +
    xlab("Cluster") + ylab("Span EMM")

#ggsave(here("output", "figures", ROUND, "emm_clusters.png"), dpi = 300,
            #units='in', width = 6, height = 2.5)
ggsave(here("output", "figures", ROUND, "emm_clusters_spinh.png"), plot=p_1, dpi = 300,
            units='in', width = 2, height = 2.5)
ggsave(here("output", "figures", ROUND, "emm_clusters_spinl.png"), plot=p_2, dpi = 300,
            units='in', width = 2, height = 2.5)
ggsave(here("output", "figures", ROUND, "emm_clusters_span.png"), plot=p_3, dpi = 300,
            units='in', width = 2, height = 2.5)

```

The above tables show an overall effect of clustering for SPiN-High and SPAN, even when controlling for demographic variables. Next, we visualize this relationship in terms of the differences in performance among clusters.

Estimated marginal means are given for models containing clusters as a predictor.
